``Interestingness'' in Mathematical Theory Exploration

### 2019-10-18 PhDThesis/litreview_staticanalysis.tex
Citations

Superoptimisers are often restricted to a branch- and effect-free subset of
machine code, where cost can be reasonably estimated by counting instructions or
cycles. Higher-level languages require more elaborate estimation methods, such
as cost semantics~\cite{danner2015denotational}: such models will still contain
many parameters, some of which may be determined by empirical measurement,
whilst others may depend on characteristics of the input data. Despite this
drawback, pareto-optimal optimisations can still be identified, and either
presented to the user or tested against a realistic data set (known as
\emph{profile-guided optimisation}~\cite{TODO}).

@@

Rather than optimising individual units of code in isolation, we can instead
work ``top down'' and perform \emph{whole program optimisation}.\iffalse TODO: Cite MLTon, Stalin,
maybe others \fi In particular, for every function we may want to optimise, we
know \emph{all} call sites of that function within the program, and hence avoid
the need for a general solution. Statically-known values, including arguments to
(higher-order) functions, can be propagated through the code using techniques
like \emph{supercompilation}~\cite{Turchin:1986:CS:5956.5957}, and constant
expressions can be ahead of time; however this can dramatically increase code
size, as each function may be duplicated into multiple specialised instances.

@@

\hsearch{} is mostly of theoretical interest, being described as ``another $O()$
warning showing, how important [to running time] factors, and even subdominant
additive terms, are.''~\cite{TODO: Hutter, page 4}. Despite this caveat, the
general framework is still useful to compare with the code improvement
techniques described above. In particular, the use of open-ended search (process
A) is similar to that of superoptimisation; although the former enumerates
equivalence proofs and extracts programs from them, whilst the latter enumerates
programs and attempts to prove (or approximately determine)
equivalence. Likewise the speed estimation of process B is analogous to those
employed by superoptimisation. JIT compilation can also be achieved within this
framework, for example by including our desired optimisations in the axioms of
the formal system (making proof search trivial), biasing the search algorithm to
consider applications of these optimisations first, and including a simple cost
model for the language. Switching to faster programs can also be made
incremental, e.g. allowing (stateless) functions to be replaced individually
without having to restart execution, just like in a JIT-enabled interpreter.

>>>>

Superoptimisers are often restricted to a branch- and effect-free subset of
machine code, where cost can be reasonably estimated by counting instructions or
cycles. Higher-level languages require more elaborate estimation methods, such
as cost semantics~\cite{danner2015denotational}: such models will still contain
many parameters, some of which may be determined by empirical measurement,
whilst others may depend on characteristics of the input data. Despite this
drawback, pareto-optimal optimisations can still be identified, and either
presented to the user or tested against a realistic data set (known as
\emph{profile-guided optimisation}~\cite{hubicka2005profile}).

@@

Rather than optimising individual units of code in isolation, we can instead
work ``top down'' and perform \emph{whole program
  optimisation}~\cite{weeks2006whole,siskind1999flow}. In particular, for every
function we may want to optimise, we know \emph{all} call sites of that function
within the program, and hence avoid the need for a general solution.
Statically-known values, including arguments to (higher-order) functions, can be
propagated through the code using techniques like
\emph{supercompilation}~\cite{Turchin:1986:CS:5956.5957}, and constant
expressions can be ahead of time; however this can dramatically increase code
size, as each function may be duplicated into multiple specialised instances.

@@

\hsearch{} is mostly of theoretical interest, being described as ``another $O()$
warning showing, how important [to running time] factors, and even subdominant
additive terms, are.''~\cite[p. 4]{HUTTER:2002}. Despite this caveat, the
general framework is still useful to compare with the code improvement
techniques described above. In particular, the use of open-ended search (process
A) is similar to that of superoptimisation; although the former enumerates
equivalence proofs and extracts programs from them, whilst the latter enumerates
programs and attempts to prove (or approximately determine)
equivalence. Likewise the speed estimation of process B is analogous to those
employed by superoptimisation. JIT compilation can also be achieved within this
framework, for example by including our desired optimisations in the axioms of
the formal system (making proof search trivial), biasing the search algorithm to
consider applications of these optimisations first, and including a simple cost
model for the language. Switching to faster programs can also be made
incremental, e.g. allowing (stateless) functions to be replaced individually
without having to restart execution, just like in a JIT-enabled interpreter.
